# Groundwater

Purpose: derive metrics of groundwater influence for reporting unit locations

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(terra)
library(knitr)
library(viridis)
```

## Data

Load spring prevalence rasters, lakes, basin and flowline
```{r}
# spring prevalence from Maxent: complete tiff and buffered/no lake tiff
spring_full <- rast("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Groundwater Seeps/SpringTIFFs/SpringPrev_SnakeGreys_BedSurf.tif")

# lakes in Upper Snake
lakes <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Lakes/UpperSnake_Lakes.shp") 

# basin without lakes
basin <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Basin Delineation/BasinDelineation/MajorBasins_Watersheds.shp") 
basin <- subset(basin, basin$site %in% c("UpperSnake", "Greys"))
basin_nolakes <- erase(basin, lakes)

# flowline
flowline <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Watershed Delineation/SnakeGreys_flowline.shp")
flowbuff <- flowline %>% buffer(width = 100)
```

Remove lakes and buffer to flowline
```{r eval=FALSE}
spring_nolakes <- mask(spring_full, basin_nolakes)
spring_nolakes_buff100 <- mask(spring_nolakes, flowbuff)
```

Write-out spring prevalence rasters
```{r eval=FALSE}
writeRaster(spring_nolakes, "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Groundwater Seeps/SpringTIFFs/SpringPrev_SnakeGreys_BedSurf_nolakes.tif")
writeRaster(spring_nolakes_buff100, "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Groundwater Seeps/SpringTIFFs/SpringPrev_SnakeGreys_BedSurf_nolakes_flowbuff100.tif")
```

Load spring prevalence rasters
```{r}
spring_nolakes <- rast("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Groundwater Seeps/SpringTIFFs/SpringPrev_SnakeGreys_BedSurf_nolakes.tif")
spring_nolakes_buff100 <- rast("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Groundwater Seeps/SpringTIFFs/SpringPrev_SnakeGreys_BedSurf_nolakes_flowbuff100.tif")
```

View MaxEnt spring prevalnce raster
```{r fig.width=4, fig.height=7}
options(terra.pal=rev(viridis(100)))
plot(spring_nolakes)
```

Buffered to flowline...
```{r fig.width=4, fig.height=7}
plot(spring_nolakes_buff100)
```

Reporting unit locations and watersheds
```{r}
# reporting unit locations
sites_repu <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Location Data/RepUnit_SpatialLocations.shp")

# watershed shapefiles
sheds_repu <- read_sf(dsn = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Watershed Delineation", layer = "RepUnits_Watersheds")
```

View reporting units and watersheds
```{r}
mapview(list(st_as_sf(sites_repu), sheds_repu), col.regions = c("blue", "white"), col = c("blue", "blue"), alpha.regions = c(1,0.2), legend = F)
```


## Derive groundwater metrics

Calculate groundwater metrics for each basin. Extract average and weighted spring prevalence for each basin
```{r eval=FALSE}
# 
sites <- sheds_repu$site
gwlist <- list()
st <- Sys.time()
for (i in 1:length(sites)) {
  spring_mask <- mask(crop(spring_nolakes_buff100, sheds_repu[sheds_repu$site == sites[i],]), sheds_repu[sheds_repu$site == sites[i],]) # crop and mask by basin
  dist_rast <- distance(spring_mask, sites_repu[sites_repu$repunit == sites[i],]) %>% mask(spring_mask) # calculate distance between each raster cell and site location
  gwlist[[i]] <- tibble(site = sites[i],
                        areasqkm = sheds_repu$aresqkm[i],
                        gwi_point = terra::extract(spring_nolakes_buff100, sites_repu[sites_repu$repunit == sites[i],], na.rm = TRUE)[,2],
                        gwi_iew05km = as.numeric(global(spring_mask * (1 / exp(dist_rast/5000)), "sum", na.rm = T) / global(1 / exp(dist_rast/5000), "sum", na.rm = T))
  )
  print(i)
}
Sys.time() - st
gwmetrics_repu <- do.call(rbind, gwlist) # bind as tibble
write_csv(gwmetrics_repu, "Landscape Covariates/Groundwater/GroundwaterMetrics_raw_RepUnits.csv")
```

```{r}
gwmetrics_repu <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Groundwater/GroundwaterMetrics_raw_RepUnits.csv")
gwmetrics_repu %>% kable()
```

```{r}
sites_repu <- sort(sites_repu, "repunit")
gwmetrics_repu <- arrange(gwmetrics_repu, site)
sites_repu$gwi <- gwmetrics_repu$gwi_iew05km
mapview(sites_repu, zcol = "gwi", col.regions = viridis::viridis(n = 10, direction = -1))
```

