# Baseline Fst

Purpose: calculate index of genetic differentiation (Fst) among baseline reporting groups. See Willing et al. (2012, PLoS One) for details on how number of SNPs and sample size (# of individuals) affects the accuracy of Fst.

```{r echo=FALSE}
library(tidyverse)
library(hierfstat)
library(plot.matrix)
library(viridis)
library(reactable)
```

## Load data

Load genotyp data, calculate reporting group sample size, and load GPS location data:
```{r}
# load genotype data
refdat <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Baseline Testing/UpperSnakeRiver_GTseq_InputData_NoSibs_clean_baseline.csv") %>% mutate_all(~replace_na(., 0))

# get sample size
ss <- refdat %>% group_by(repunit) %>% summarize(ss = n())

# location data for sorting by latitude
gps <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/RepUnit_LatLong.csv") %>% arrange(desc(lat))
```

Reformat genotype data for hierfstat:
```{r}
refdat_sub <- refdat[,c(5:ncol(refdat))] %>% unite(vnew, sep = "") %>% 
  separate(vnew, into = colnames(refdat)[c(seq(from = 5, to = ncol(refdat), by = 2))], sep = c(seq(from = 2, to = 532, by = 2))) 
refdat_sub <- as_tibble(data.frame(sapply(refdat_sub, as.numeric))) %>% mutate_all(~na_if(., 0))
refdat_sub$repunit <- refdat$repunit
refdat_sub <- refdat_sub %>% relocate(repunit)
```

Calculate basic statistics:
```{r}
mybs <- basic.stats(data.frame(refdat_sub))
print(mybs)
```


## Pairwise Fst

Calculate pairwise Fst following Weir and Cockerham (1984)...also see Harris et al (2022) Conservation Genetics. NOTE: this takes ~1 hr to run
```{r}
# ref_fst <- pairwise.WCfst(data.frame(refdat_sub))
# write.csv(data.frame(ref_fst), "GSI Analysis/Baseline Fst/UpperSnakeRiver_GTseq_BaselinePopsFst.csv", row.names = T)
ref_fst <- read.csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Baseline Fst/UpperSnakeRiver_GTseq_BaselinePopsFst.csv")
ref_fst_bylat <- tibble(ref_fst[,c("X", gps$repunit)]) %>% arrange(match(X, gps$repunit)) %>% column_to_rownames("X")
reactable(round(ref_fst_bylat, digits = 3))
```

Range of among-population Fst:
```{r}
range(ref_fst_bylat, na.rm = T)
```

Plot histogram of pairwise Fst values. Vertical red line denotes Fst = 0.01: the recommended minimum pairwise Fst for accurate genetic stock identification (Araujo et al. 2014, TAFS). Note that most pariwise comparisons (among reporting groups) are above this threshold.
```{r}
par(mar = c(4,4,1,1))
hist(ref_fst_bylat[lower.tri(ref_fst_bylat)], xlab = expression("F"["st"]), main = "", breaks = 50)
abline(v = 0.01, col = "red", lwd = 1.5)
```
```{r echo=FALSE, message=FALSE, results='hide'}
jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Baseline Fst/BaselinePopsFst_hist.jpg", res = 1000, units = "in", width = 4.5, height = 4.25)
par(mar = c(4,4,1,1))
hist(ref_fst_bylat[lower.tri(ref_fst_bylat)], xlab = expression("F"["st"]), main = "", breaks = 50)
abline(v = 0.01, col = "red", lwd = 1.5)
dev.off()
```

Proportion of reporting group pairs for which Fst < 0.01
```{r}
rrr <- ref_fst_bylat[lower.tri(ref_fst_bylat)]
length(rrr[rrr < 0.01]) / length(rrr)
```

For which pairs is Fst < 0.01?

* Fst indicates low differentiation/lots of mixing among UBBC, Cowboy Cabin, Snake River Side Channel, and Blacktail.
```{r}
fstlist <- lapply(apply(ref_fst_bylat < 0.01, 1, which), names)
fstlist[lengths(fstlist) > 0]
```

Plot matrix of pairwise Fst: 
```{r fig.width=12, fig.height=12}
# plot color matrix of Fst
r2 <- ref_fst_bylat
r2[upper.tri(r2)] <- NA

par(mar = c(12,12,4,4))
plot(as.matrix(r2), main = expression("F"["st"]), col = rev(magma(20)), las = 2, xlab = "", ylab = "", axes = F, border = "grey50")
axis(1, cex.axis = 0.2)
axis(2, cex.axis = 0.2)
```
```{r echo=FALSE, message=FALSE, results='hide'}
jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Baseline Fst/BaselinePopsFst_colormatrix_byLat.jpg", res = 1000, units = "in", width = 12, height = 12)
par(mar = c(12,12,4,4))
plot(as.matrix(r2), main = expression("F"["st"]), col = rev(magma(20)), las = 2, xlab = "", ylab = "", axes = F, border = "grey50")
axis(1, cex.axis = 0.2)
axis(2, cex.axis = 0.2)
dev.off()
```

The same plot, but with numbered reporting groups (for supporting information):
```{r fig.width=10, fig.height=10}
# plot color matrix of Fst
rownames(r2) <- 1:dim(r2)[1]
colnames(r2) <- 1:dim(r2)[1]

par(mar = c(2.5,2.5,1,4))
plot(as.matrix(r2), main = "", col = rev(magma(20)), las = 2, xlab = "", ylab = "", axes = F, border = "grey50")
axis(1, cex.axis = 0.2)
axis(2, cex.axis = 0.2)
```
```{r}
jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Baseline Fst/BaselinePopsFst_colormatrix_byLat_numbered.jpg", res = 1000, units = "in", width = 10, height = 10)
par(mar = c(2.5,2.5,4,4))
plot(as.matrix(r2), main = expression("F"["st"]), col = rev(magma(20)), las = 2, xlab = "", ylab = "", axes = F, border = "grey50")
axis(1, cex.axis = 0.2)
axis(2, cex.axis = 0.2)
dev.off()
```


For Wy-ACT Fish-Flows manuscript, get range and median Fst among spring creeks with redd count data
```{r}
mypops <-  c("blacktail_NA", "cody_bluecrane", "cowboycabin_NA", "fish_NA", "flat_NA", "lowerbarbc_NA", "snakeriversidechannel_NA", "spring_tss", "threechannel_NA", "upperbarbc_NA")
ref_fst2 <- ref_fst_bylat[mypops, mypops]
ref_fst3 <- ref_fst_bylat[lower.tri(ref_fst2)]
range(ref_fst3, na.rm = T)
median(ref_fst3, na.rm = T)
```

