---
title: "Obj. 1 - Spatiotemporal variation"
---

Purpose: quantify the degree to which tributary contributions to the mainstem Snake River population of YCT change over space and/or time.

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(zoid)
library(loo)
library(rstan)
library(ggfortify)
library(robCompositions)
library(knitr)
```


```{r include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, cache.lazy = FALSE)
```


## Data

Load GSI output from rubias
```{r}
gsi_dat <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/GSI Analysis/By Section and Year/UpperSnake_GSI_SectionYear_output.RDS")
```

Create data frame of bootstrap-corrected mixing abundance by multiplying bootstrap-corrected mixing proportions by section and year-specific sample size. Because we used the bootstrap approach to correct mixing proportions by estimated bias in reporting unit contributions (see Chapters 3 and 5), we cannot just sum the individual posteriors as in Jensen *et al.* (2022). Instead, simply multiply bootstrap-corrected mixing proportions by compsition sample size to get corrected abundance. This is functionally the same as the approach in Jensen *et al.* (2022).
```{r}
# abundance for zoid
gsi_bscorr <- gsi_dat$bootstrapped_proportions %>%  
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>% 
  mutate(bscorrnum = bs_corrected_repunit_ppn*sampsize, section = str_sub(mixture_collection, end = -6), year = str_sub(mixture_collection, start = -4)) %>%
  select(section, year, repunit, bscorrnum) %>% spread(key = repunit, value = bscorrnum)
gsi_bscorr

# proportions for PCA viz
gsi_bscorr2 <- gsi_dat$bootstrapped_proportions %>%  
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>% 
  mutate(bscorrnum = bs_corrected_repunit_ppn, section = str_sub(mixture_collection, end = -6), year = str_sub(mixture_collection, start = -4)) %>%
  select(section, year, repunit, bscorrnum) %>% spread(key = repunit, value = bscorrnum) %>%
  mutate(section2 = factor(recode(section, 
                            "snake_pacificdeadmans" = "A",
                            "snake_deadmansmoose" = "B",
                            "snake_moosewilson" = "C",
                            "snake_wilsonsouthpark" = "D",
                            "snake_southparkastoria" = "E",
                            "snake_astoriawesttable" = "F"), levels = c("A", "B", "C", "D", "E", "F"))) %>%
  relocate(section2, .after = section)
```

Construct data (bootstrapped mixing abundance) and design (section, year) matrics for zoid
```{r}
data_matrix <- gsi_bscorr[,-c(1:2)]
design_matrix <- gsi_bscorr[,c(1,2)]
design_matrix$section <- as.factor(design_matrix$section)
design_matrix$year <- as.factor(design_matrix$year)
design_matrix$y <- 1 # dummy variable as in Jensen et al. 2022 and zoid package vignette
```

### Visualize data with PCA

Run PCA on bootstrap-corrected proportional contribution. Not recommended to run PCA on proportional data without transformation, but we do this here just to visualize general patterns in the data
```{r}
# run PCA
gsi_pca <- prcomp(gsi_bscorr2[,-c(1:3)])
# view big drivers
gsi_pca$rotation[gsi_pca$rotation[,1] > 0.2 | gsi_pca$rotation[,1] < -0.2 | gsi_pca$rotation[,2] > 0.2 | gsi_pca$rotation[,2] < -0.2 ,c(1:2)]
```
Plot, trim to only biggest loadings: lots of clustering by section, not so much by year
```{r}
p0 <- autoplot(gsi_pca, data = gsi_bscorr2, colour = "section2", shape = "year", loadings = TRUE, loadings.label = TRUE, loadings.colour = "black", size = 3) 
p0$layers[[2]]$data <- p0$layers[[2]]$data[c("bailey_NA", "blacktail_NA", "cabin_NA", "fish_NA", "mosquito_NA", "upperbarbc_NA", "willow_NA"),]
p0$layers[[3]]$data <- p0$layers[[3]]$data[c("bailey_NA", "blacktail_NA", "cabin_NA", "fish_NA", "mosquito_NA", "upperbarbc_NA", "willow_NA"),]
p0 + scale_color_manual(values = hcl.colors(6, "Zissou 1"))
```

As expected, sections clump together with no major pattern by year. Contributions from Upper Bar BC and Blacktail are characteristic of sections A and B (upstream); Fish and Mosquito of section D; and Bailey, Cabin and Willow of sections E and F


## Model using ZOID

Fit candidate models
```{r eval=FALSE}
# Section * Year
fitz1 <- fit_zoid(formula = y ~ section * year, design_matrix = design_matrix, data_matrix = as.matrix(data_matrix), overdispersion = FALSE, chains = 3, iter = 4000, posterior_predict = TRUE, moment_match = TRUE)

# Section + Year
fitz2 <- fit_zoid(formula = y ~ section + year, design_matrix = design_matrix, data_matrix = as.matrix(data_matrix), overdispersion = FALSE, chains = 3, iter = 4000, posterior_predict = TRUE, moment_match = TRUE)

# Section
fitz3 <- fit_zoid(formula = y ~ section, design_matrix = design_matrix, data_matrix = as.matrix(data_matrix), overdispersion = FALSE, chains = 3, iter = 4000, posterior_predict = TRUE, moment_match = TRUE)

# Year
fitz4 <- fit_zoid(formula = y ~ year, design_matrix = design_matrix, data_matrix = as.matrix(data_matrix), overdispersion = FALSE, chains = 3, iter = 4000, posterior_predict = TRUE, moment_match = TRUE)

# Null
fitz5 <- fit_zoid(formula = y ~ 1, design_matrix = design_matrix, data_matrix = as.matrix(data_matrix), overdispersion = FALSE, chains = 3, iter = 4000, posterior_predict = TRUE, moment_match = TRUE)

# save model objects
saveRDS(fitz1, file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_1.rds")
saveRDS(fitz2, file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_2.rds")
saveRDS(fitz3, file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_3.rds")
saveRDS(fitz4, file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_4.rds")
saveRDS(fitz5, file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_5.rds")
```

Load fitted model objects
```{r cache=FALSE}
fitz1 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_1.rds")
fitz2 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_2.rds")
fitz3 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_3.rds")
fitz4 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_4.rds")
fitz5 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/Candidate model outputs/ZOID_CandidateModel_5.rds")
```


### Model selection
```{r}
fit_list <- list(fitz1, fitz2, fitz3, fitz4, fitz5)
loo_list <- list()
for (i in 1:length(fit_list)) {
  m <- fit_list[[i]]
  loo_list[[i]] <- loo::loo(m$model)
  #print(i)
}
lc <- loo_compare(loo_list)
print(lc, simplify = FALSE, digits = 2)
plot(loo_list[[3]])
write.csv(as.data.frame(lc), "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/ZOID_ModelSelection_LOOCV_Output.csv")

# set top model
top_mod <- fitz3
```

### PP check

Join posterior predicted proportional contributions to original data
```{r}
# get model predictioned proportions
preds <- get_fitted(top_mod)

# join to observed proportions
obspred <- gsi_bscorr2 %>% gather(bailey_NA:willow_NA, key = "repunit", value = "bs_ppn") %>%
  mutate(pred_mean = preds$mean,
         pred_median = preds$median,
         pred_lo = preds$lo,
         pred_hi = preds$hi)
```

Plot predictions by observed, all together, shape by year and color by section. Note that the 0's are not very well predicted, which may drive lower than observed predictions from reporting groups contributing a lot to the Snake. 
```{r  fig.width=5, fig.height=4.3}
obspred %>%
  ggplot(aes(x = bs_ppn, y = pred_median)) +
  geom_point(aes(color = section2, shape = year)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  scale_color_manual(values = hcl.colors(6, "Zissou 1")) +
  xlab("Observed mixing proportion") + ylab("Predicted mixing proportion") +
  theme_bw()
```

Plot predictions by observed, all together, shape by year and color by reporting unit
```{r  fig.width=5, fig.height=4.3}
obspred %>%
  ggplot(aes(x = bs_ppn, y = pred_median)) +
  geom_point(aes(color = repunit, shape = year)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  #scale_color_manual(values = hcl.colors(6, "Zissou 1")) +
  xlab("Observed mixing proportion") + ylab("Predicted mixing proportion") +
  theme_bw() + theme(legend.position = "none")
```

Plot predictions by observed, facet by year and section
```{r  fig.width=6, fig.height=9}
obspred %>%
  ggplot(aes(x = bs_ppn, y = pred_median)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, col = "red") + 
  facet_grid(section2 ~ year) +
  xlab("Observed mixing proportion") + ylab("Predicted mixing proportion") +
  theme_bw()
```


### Plot model output

Organize data and ensure proportions sum to 1 for each section
```{r}
gps <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Location Data/RepUnit_LatLong.csv")

# summarize data
fit_vals <- get_fitted(top_mod)
mixdes <- design_matrix %>% select(section, year) %>% mutate(obs = c(1:17))
tribdes <- tibble(group = 1:52, repunit = names(data_matrix))
fit_sum <- fit_vals %>% left_join(mixdes) %>% left_join(tribdes) %>% group_by(section, repunit) %>% summarize(mean = unique(mean), median = unique(median), lo = unique(lo), hi = unique(hi)) %>% ungroup() %>% left_join(gps)

# check sum to 1
fit_sum %>% group_by(section) %>% summarize(tot = sum(mean)) # check: these should sum to 1
```

Plot posterior predicted proportional contribution by reporting groups, across sections. Compare to figures in *Chapter 5 - GSI Analysis*. Note that all we are doing here is estimating section-level means!
```{r  fig.width=8, fig.height=7}
mylabs <- tibble(section2 = c("A", "B", "C", "D", "E", "F"),
                 mylab = c("A", "B", "C", "D", "E", "F"))

fit_sum %>% 
  mutate(section2 = factor(recode(section, 
                            "snake_pacificdeadmans" = "A",
                            "snake_deadmansmoose" = "B",
                            "snake_moosewilson" = "C",
                            "snake_wilsonsouthpark" = "D",
                            "snake_southparkastoria" = "E",
                            "snake_astoriawesttable" = "F"), levels = c("A", "B", "C", "D", "E", "F"))) %>%
  ggplot(aes(x = reorder(repunit, -lat), y = median)) +
  geom_bar(aes(x = reorder(repunit, -lat), y = median), stat = "identity", color = "black", fill = "grey", width = 1) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0) +
  facet_wrap(~ section2, nrow = 6) +
  xlab("Reporting group") + ylab("Predicted mixture proportion") + 
  theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5), strip.text.x = element_blank()) + 
  scale_x_discrete(labels = c(1:52)) + 
  geom_text(data = mylabs, aes(x = -Inf, y = Inf, label = mylab, hjust = -1, vjust = 1.5), size = 5)
```

