---
title: "Spatial Pattern"
---

Purpose: conduct post-hoc analyses to explore spatial structure and tributary random intercepts (mean contribution after account for the effects of distance, size, groundwater, and connectivity).

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(SSNbler)
library(SSN2)
library(brms)
library(bayesplot)
library(ggpubr)
library(tidybayes)
library(gpkg)
library(RColorBrewer)
```


## Data and objects

Load spatial datasets. Check for geometry type SSNbler/SSN2 compatibility (linestring, point) and matching CRS. 

Upper Snake-Greys-Salt Rivers flowline object
```{r}
flowline <- st_read(dsn = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Flowline/FlowlineEditing/SnakeHeadwaters_flowline_springsclean_fixed.gpkg", layer = "flowline")
flowline <- flowline[!st_is_empty(flowline),]
flowline <- flowline %>% st_cast("LINESTRING")
flowline
```

Reporting unit locations
```{r}
sites_repu <- read_sf("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Location Data/RepUnit_SpatialLocations.shp")
#sites_repu <- sort(sites_repu, "repunit")

# define spatial clusters/regions
greys <- c("littlegreys_steer", "white_NA", "blindbull_NA", "deadman_greys")
hoback <- c("boulder_NA", "granite_lower", "granite_upper", "shoal_NA", "cliff_NA", "dell_NA", "hoback_upper")
grosventre <- c("crystal_lower", "crystal_upper", "slate_NA", "goosewing_NA", "cottonwood_grosventre", "fish_grosventre")
spread <- c("rock_NA", "spread_uppermainstem", "spreadnf_flagstaff", "spread_southfork", "leidy_NA")
buffalo <- c("blackrock_lower", "blackrock_upper", "lava_NA", "box_NA", "clear_NA", "northbuffalofork_NA", "pacific_NA", "spring_nps")
lowsnake <- c("horse_NA", "fall_coburn", "willow_NA", "dog_NA", "cabin_NA", "bailey_NA")
gtnp <- c("snakeriversidechannel_NA", "upperbarbc_NA", "cottonwood_NPS", "blacktail_NA", "ditch_NA", "cowboycabin_NA", "deadmansbar_NA")

# load loation data and edit names, regions
gps <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Location Data/RepUnit_LatLong.csv") %>% 
  arrange(desc(lat)) %>%
  mutate(repunit_num = c(1:52)) %>%
  mutate(repunit2 = gsub("_NA", "", repunit)) %>%
  mutate(repunit2 = recode(repunit2, "snakeriversidechannel" = "srsidechannel", "spread_uppermainstem" = "spread_mainstem", "cottonwood_grosventre" = "cottonwood_gv")) %>%
  mutate(name = paste(repunit_num, ". ", repunit2, sep = ""),
         region = ifelse(repunit %in% greys, "Greys",
                         ifelse(repunit %in% hoback, "Hoback",
                                ifelse(repunit %in% grosventre, "Gros Ventre",
                                       ifelse(repunit %in% spread, "Spread",
                                              ifelse(repunit %in% buffalo, "Buffalo Fork", 
                                                     ifelse(repunit %in% lowsnake, "Lower Snake", 
                                                            ifelse(repunit %in% gtnp, "Grand Teton", "Jackson"))))))))
```

ZOIB model object
```{r}
mod_zoib1 <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/brms_fit_zoib_fullmodel.rds")
```

Join spatial points with model output
```{r}
rpint_summ <- mod_zoib1 %>%
  spread_draws(r_repunit[repunit,]) %>%
  group_by(repunit) %>%
  median_qi(r_repunit) %>%
  ungroup() %>%
  mutate(repunit = reorder(repunit, r_repunit)) %>%
  rename("lower" = ".lower", "upper" = ".upper", "width" = ".width", "interval" = ".interval")

sites_repu <- sites_repu %>% left_join(gps) %>% left_join(rpint_summ)
sites_repu
```

## Vizualize data

Dot plot of reporting unit random effects
```{r  fig.width=6, fig.height=8}
mod_zoib1 %>%
  spread_draws(r_repunit[repunit,]) %>%
  #compare_levels(r_repunit, by = repunit) %>%
  ungroup() %>%
  left_join(gps %>% select(repunit, name, lat, region)) %>%
  mutate(name = reorder(name, r_repunit)) %>%
  ggplot(aes(y = name, x = r_repunit)) +
  stat_pointinterval(.width = c(.80, .95)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ylab("Reporting unit") + xlab("Offset to intercept") + 
  #scale_colour_brewer(palette = "Dark2") +
  theme_bw()
```

Map flowline and reporting unit locations colored by random effect value
```{r fig.width = 5, fig.height = 8}
# ggplot() + geom_sf(data = flowline) + geom_sf(data = sites_repu, color = "purple") + coord_sf(datum = st_crs(flowline)) + theme_bw()
mapview(flowline) + mapview(sites_repu, zcol = "r_repunit", col.regions = rev(hcl.colors(10, "Blue-Red")), alpha.regions = 1)
```

Create static maps to export
```{r fig.width=6, fig.height=9}
# lakes
lakes <- read_sf("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Lakes/UpperSnake_Lakes.shp") 
lakes2 <- read_sf("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Lakes/SSB_Lakes/SSB_Lakes.shp") 
lakes2 <- subset(lakes2, lakes2$GNIS_Name == "Palisades Reservoir")
lakes2 <- st_transform(lakes2, crs = st_crs(lakes))
# lakes <- st_union(lakes, lakes2)

# order and colors
regions <- c("Snake Corridor", "Buffalo Fork", "Spread", "Gros Ventre", "Lower Snake", "Hoback", "Greys")
jittersites <- c("snakeriversidechannel_NA", "upperbarbc_NA", "cottonwood_NPS", 
                 "blacktail_NA", "threechannel_NA", "lowerbarbc_NA")
mycols <- c("grey", brewer.pal(n = 6, "Set2"))

# edit geometry to avoid overlapping points
sites_repu_edit <- sites_repu
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "snakeriversidechannel_NA", st_geometry(sites_repu %>% filter(repunit == "snakeriversidechannel_NA")) + c(0,800), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "blacktail_NA", st_geometry(sites_repu %>% filter(repunit == "blacktail_NA")) + c(0,-500), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "upperbarbc_NA", st_geometry(sites_repu %>% filter(repunit == "upperbarbc_NA")) + c(2000,0), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "cottonwood_nps", st_geometry(sites_repu %>% filter(repunit == "cottonwood_nps")) + c(-1000,0), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "threechannel_NA", st_geometry(sites_repu %>% filter(repunit == "threechannel_NA")) + c(0,2000), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "fish_grosventre", st_geometry(sites_repu %>% filter(repunit == "fish_grosventre")) + c(1000,0), geometry)))
sites_repu_edit <- sites_repu_edit %>% mutate(geometry = st_sfc(ifelse(repunit == "fish_NA", st_geometry(sites_repu %>% filter(repunit == "fish_NA")) + c(0,1000), geometry)))
# reset CRS
st_crs(sites_repu_edit) <- st_crs(sites_repu)

# point borders colors by region
p1 <- ggplot() + 
  geom_sf(data = flowline, color = "grey10") + 
  geom_sf(data = lakes, color = "grey10", fill = "grey30") + 
  geom_sf(data = lakes2, color = "grey10", fill = "grey30") + 
  geom_sf(data = sites_repu_edit, aes(fill = r_repunit, color = factor(region, levels = regions)), shape = 21, size = 7, stroke = 1.5) + 
  #coord_sf(datum = st_crs(flowline)) + 
  scale_color_manual(values = mycols) + 
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  xlim(510000,567000) + ylim(4757000,4857000) +
  theme_bw() + 
  theme(legend.position = "none", panel.background = element_rect(fill = "grey50"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# no color by region
p2 <- ggplot() + 
  geom_sf(data = flowline, color = "grey10") + 
  geom_sf(data = lakes, color = "grey10", fill = "grey30") + 
  geom_sf(data = lakes2, color = "grey10", fill = "grey30") + 
  geom_sf(data = sites_repu_edit, aes(fill = r_repunit), shape = 21, size = 7) + 
  #coord_sf(datum = st_crs(flowline)) + 
  coord_sf(label_axes = list(
      top = "E",
      left = "N"
    )) +
  scale_color_manual(values = mycols) + 
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  xlim(510000,567000) + ylim(4757000,4857000) +
  theme_bw() + 
  theme(legend.position = "none", panel.background = element_rect(fill = "grey50"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1))
```

Save maps to file
```{r}
jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/RandomIntercepts_Map_Outlined.jpg", height = 8.5, width = 5.5, units = "in", res = 1000)
p1 
dev.off()

jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/RandomIntercepts_Map.jpg", height = 8.5, width = 5.5, units = "in", res = 1000)
p2
dev.off()
```


## Build the LSN

Specify LSN file path
```{r}
lsn.path <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/lsn"
```


Convert linestring to landscape network. Ensure 0 topology errors, 1 outlet!
```{r}
edges <- lines_to_lsn(
  streams = flowline,
  lsn_path = lsn.path,
  check_topology = TRUE,
  snap_tolerance = 0.05,
  topo_tolerance = 20,
  overwrite = TRUE
)
```

Load edges and nodes
```{r}
edges <- st_read(dsn = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/lsn/edges.gpkg", layer = "edges")
nodes <- st_read(dsn = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/lsn/nodes.gpkg", layer = "nodes")
```

View outlet
```{r}
mapview(edges, zcol = "FID_1") + mapview(nodes %>% filter(nodecat == "Outlet"), zcol = "nodecat")
```

View confluences
```{r}
mapview(edges, zcol = "FID_1") + mapview(nodes %>% filter(nodecat == "Confluence"), zcol = "nodecat")
```

View confluences
```{r}
mapview(edges, zcol = "FID_1") + mapview(nodes %>% filter(nodecat == "Source"), zcol = "nodecat")
```


## Add sites to LSN

```{r}
obs <- sites_to_lsn(
  sites = sites_repu,
  edges = edges,
  lsn_path = lsn.path,
  file_name = "obs",
  snap_tolerance = 200,
  save_local = TRUE,
  overwrite = TRUE
)
```

Check snapping
```{r}
mapview(edges) + mapview(sites_repu, col.regions = "black") + mapview(obs, col.regions = "red")
```


## Upstream distance

Calculate upstream distance (from outlet) for edges. Not sure if this step is necessary for Torgegram. 
```{r}
edges <- updist_edges(
  edges = edges,
  save_local = TRUE,
  lsn_path = lsn.path,
  calc_length = TRUE
)
names(edges) ## View edges column names
```

Update sites with upstream distance
```{r}
site.list <- updist_sites(
  sites = list(
    obs = obs
  ),
  edges = edges,
  length_col = "Length",
  save_local = TRUE,
  lsn_path = lsn.path
)
names(site.list$obs) ## View output site.list names
```

View upstream distance for edges and sites
```{r}
mapview(edges, zcol = "upDist") + mapview(site.list$obs, zcol = "upDist")
```


## Assemble SSN

Assemble SSN object, no prediction sites, no AFVs
```{r}
snake_ssn <- ssn_assemble(
  edges = edges,
  lsn_path = lsn.path,
  obs_sites = site.list$obs,
  #preds_list = site.list[c("pred1km", "CapeHorn")],
  ssn_path = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/SnakeHeadwaters.ssn",
  import = TRUE,
  check = TRUE,
  #afv_col = "afvArea",
  overwrite = TRUE
)
```

View object details
```{r}
snake_ssn
```

Check SSN class, names
```{r}
class(snake_ssn) ## Get class
names(snake_ssn) ## print names of SSN object
names(snake_ssn$obs) ## print names of observation variables
```

Plot the SSN object
```{r}
ggplot() +
  geom_sf(
    data = snake_ssn$edges,
    color = "medium blue",
  ) +
  scale_linewidth(range = c(0.1, 2.5)) +
  geom_sf(
    data = snake_ssn$obs,
    size = 1.7,
    aes(color = r_repunit)
  ) +
  scale_color_viridis_c()
```

Supplement SSN object with hydrologic distance matrices that preserve flow directionality. 
```{r}
ssn_create_distmat(
  ssn.object = snake_ssn,
  overwrite = TRUE
)
```


## Torgegram

Plot a Torgegram (Zimmerman and Ver Hoef, 2017) to understand the type and degree of spatial structure in reporting unit random intercepts (i.e., mean contribution from each reporting unit after accounting for the effects of distance, catchment area, groundwater, and connectivity). Interpret Torgram visually sensu McGuire *et al.* (2014, PNAS) and Brennan *et al.* (2016, Geophysical Research Letters)

Compute the Torgegram:
```{r}
tg <- Torgegram(
  formula = r_repunit ~ 1,
  ssn.object = snake_ssn,
  type = c("flowuncon", "euclid"),
  bins = 20
)
tg
```

Plot torgegram
```{r}
torg_tib <- bind_rows(tg$flowuncon %>% mutate(type = "Watercourse"),
                      tg$euclid %>% mutate(type = "Euclidean")) %>%
  mutate(distkm = dist/1000)
torg_tib %>% ggplot(aes(x = dist/1000, y = gamma, group = type, color = type)) + 
  geom_smooth(se = FALSE, span = 0.7) + 
  scale_color_manual(values = c("darkorange", "dodgerblue")) +
  geom_point(aes(fill = type, size = np), color = "black", shape = 21) +
  scale_fill_manual(values = c("darkorange", "dodgerblue")) +
  xlab("Distance (kms)") + ylab("Semivariance") + #ylim(0.19,0.61) +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Pretty plot
```{r}
# convert to kms
tg$flowuncon$distkm <- tg$flowuncon$dist/1000
tg$euclid$distkm <- tg$euclid$dist/1000

# plot
par(mar = c(4,4,1,1), mgp = c(2.5,1,0))
plot(gamma ~ distkm, torg_tib, type = "n", xlab = "Distance (kms)", ylab = "Semivariance", bty = "l")
# euclidean distance
points(gamma ~ distkm, torg_tib %>% filter(type == "Euclidean"), pch = 21, bg = "black")
fit <- loess(gamma ~ distkm, tg$euclid[-2,], span = 0.7)
j <- order(tg$euclid[-2,]$distkm)
lines(tg$euclid[-2,]$distkm[j], fit$fitted[j], col="black",lwd = 2)
arrows(x0 = 15, x1 = 15, y0 = 0, y1 = 0.355, lty = 3, lwd = 2, col = "black", length = 0)
# watercourse distance
points(gamma ~ distkm, torg_tib %>% filter(type == "Watercourse"), pch = 21, bg = "grey70")
fit <- loess(gamma ~ distkm, tg$flowuncon, span = 0.7)
j <- order(tg$flowuncon$distkm)
lines(tg$flowuncon$distkm[j], fit$fitted[j], col="grey70",lwd = 2)
arrows(x0 = 44, x1 = 44, y0 = 0, y1 = 0.48, lty = 3, lwd = 2, col = "grey70", length = 0)
# legend
legend("topright", legend = c("Euclidean", "Watercourse"), fill = c("black", "grey70"), bty = "n")

```

```{r}
jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/RandomIntercepts_Torgegram.jpg", height = 3.75, width = 4, units = "in", res = 1000)

# plot
par(mar = c(3.5,3.5,0.5,0.1), mgp = c(2,0.8,0))
plot(gamma ~ distkm, torg_tib, type = "n", xlab = "Distance (km)", ylab = "Semivariance", bty = "l")
# euclidean distance
fit <- loess(gamma ~ distkm, tg$euclid[-2,], span = 0.7)
j <- order(tg$euclid[-2,]$distkm)
lines(tg$euclid[-2,]$distkm[j], fit$fitted[j], col="black", lwd = 3)
points(gamma ~ distkm, torg_tib %>% filter(type == "Euclidean"), pch = 21, bg = "black", col = "grey60")
# watercourse distance
fit <- loess(gamma ~ distkm, tg$flowuncon, span = 0.7)
j <- order(tg$flowuncon$distkm)
lines(tg$flowuncon$distkm[j], fit$fitted[j], col="grey60",lwd = 3)
points(gamma ~ distkm, torg_tib %>% filter(type == "Watercourse"), pch = 21, bg = "grey70")
# mark inflection points
arrows(x0 = 15.5, x1 = 15.5, y0 = 0, y1 = 0.355, lty = 3, lwd = 2, col = "black", length = 0)
arrows(x0 = 44, x1 = 44, y0 = 0, y1 = 0.48, lty = 3, lwd = 2, col = "grey60", length = 0)
# legend
legend("topright", legend = c("Euclidean", "Watercourse"), fill = c("black", "grey70"), bty = "n")
dev.off()
```


## Basin comparison

Organize data
```{r}
# order and colors
regions <- c("Buffalo Fork", "Spread", "Grand Teton", "Gros Ventre", "Jackson", "Lower Snake", "Hoback", "Greys")
region_short <- c("BF", "SP", "GT", "GV", "JK", "LS", "HB", "GR")
mycols <- c("grey", brewer.pal(n = 7, "Set2"))

coltib <- tibble(region = factor(regions, levels = regions), cols = mycols, region_short = factor(region_short, levels = region_short))

# pull data
repunittib <- tibble(sites_repu) %>%
  mutate(region = factor(region, levels = regions)) %>%
  left_join(coltib) %>%
  arrange(region)
```

Variances equal? Differences among groups?
```{r}
bartlett.test(r_repunit ~ region, repunittib)
oneway.test(r_repunit ~ region, repunittib, var.equal = FALSE)
```

Plot
```{r}
# plot
par(mar = c(5,4,1,1), mgp = c(2.5,1,0))
boxplot(r_repunit ~ region, repunittib, col = mycols, xlab = "", ylab = expression(paste("Reporting unit random intercept (", alpha, "'"[r], ")")), axes = FALSE)
axis(1, at = c(1:8), labels = FALSE)
text(seq(1, 8, by = 1), par("usr")[3] - 0.2, labels = regions, srt = 20, pos = 1, xpd = TRUE)
axis(2)
mtext("Region", side = 1, line = 3.5)
stripchart(r_repunit ~ region, repunittib, method = "jitter", pch = 1, add = TRUE, vertical = TRUE)
```

```{r}

jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/RandomIntercepts_Boxplot.jpg", height = 3.75, width = 4, units = "in", res = 1000)
par(mar = c(3.5,3.5,0.5,0.1), mgp = c(2,0.8,0))
# boxplot(r_repunit ~ region, repunittib, col = mycols, xlab = "", ylab = expression(paste("Reporting unit random intercept (", alpha, "'"[r], ")")), axes = FALSE)
# axis(1, labels = FALSE)
# text(seq(1, 7, by = 1)-0.3, par("usr")[3] - 0.3, labels = regions, srt = 30, pos = 1, xpd = TRUE)
# axis(2)
# box(bty = "l")
# mtext("Region", side = 1, line = 4)
# stripchart(r_repunit ~ region, repunittib, method = "jitter", pch = 16, cex = 0.7, add = TRUE, vertical = TRUE)

boxplot(r_repunit ~ region_short, repunittib, col = mycols, xlab = "Region", ylab = expression(paste("Reporting unit random intercept (", alpha, "'"[r], ")")), axes = F)
axis(1, at = c(1:8), labels = region_short)
axis(2)
# text(seq(1, 8, by = 1)-0.5, par("usr")[3] - 0.27, labels = regions, srt = 40, pos = 1, xpd = TRUE)
# axis(2)
box(bty = "l")
#mtext("Region", side = 1, line = 4)
stripchart(r_repunit ~ region_short, repunittib, method = "jitter", pch = 16, cex = 0.7, add = TRUE, vertical = TRUE)

dev.off()
```


Is there a relationship between random intercept and latitude?
```{r}
repunittib %>% ggplot(aes(x = lat, y = r_repunit)) + geom_point() + geom_smooth(method = "lm")
mod <- lm(r_repunit ~ lat, repunittib)
summary(mod)
```


```{r}
x <- seq(from = min(repunittib$lat), to = max(repunittib$lat), length.out = 100)
preds <- predict(mod, newdata = list(lat = x), se.fit = TRUE, interval = "confidence", level = 0.95)

# order and colors
# coltib <- tibble(region = regions, cols = mycols, region_short = c("BF", "SP", "GT", "JK", "LS", "HB", "GR"))
# repunittib <- repunittib %>% left_join(coltib)

jpeg("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Spatial Pattern/RandomIntercepts_IntxLat.jpg", height = 3.75, width = 3.5, units = "in", res = 1000)
par(mar = c(3.5,3.5,0.5,0.1), mgp = c(2,0.8,0))
plot(r_repunit ~ lat, repunittib,type = "n", xlab = "Latitude", ylab = expression(paste("Reporting unit random intercept (", alpha, "'"[r], ")")), bty = "l", xlim = c(max(repunittib$lat), min(repunittib$lat)))
polygon(x = c(x, rev(x)), y = c(preds$fit[,2], rev(preds$fit[,3])), border = NA, col = "grey")
points(r_repunit ~ lat, repunittib, pch = 21, bg = repunittib$cols)
lines(preds$fit[,1] ~ x, lwd = 3)
dev.off()
```


```{r fig.width=5, fig.height=5}
par(mar = c(3.5,3.5,0.5,0.1), mgp = c(2,0.8,0))
plot(r_repunit ~ lat, repunittib,type = "n", xlab = "Latitude", ylab = expression(paste("Reporting unit random intercept (", alpha, "'"[r], ")")), bty = "l", xlim = c(max(repunittib$lat), min(repunittib$lat)))
polygon(x = c(x, rev(x)), y = c(preds$fit[,2], rev(preds$fit[,3])), border = NA, col = "grey")
points(r_repunit ~ lat, repunittib, pch = 21, bg = repunittib$cols)
lines(preds$fit[,1] ~ x, lwd = 3)
```

