# Obj 1 - Spatiotemporal Variability

Purpose: assess the degree to which tributary contributions to the Snake River population of Yellowstone cutthroat trout vary across space and time

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(scales)
library(R2jags)
library(MCMCvis)
library(loo)
library(viridis)
library(beepr)
library(GGally)
library(ggmcmc)
library(zoid)
library(brms)
library(bayesplot)
library(ggpubr)
library(tidybayes)
```

## Data

### Bootstrapped mixing proportions
```{r}
gsi_dat <- readRDS("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/GSI Analysis/By Section and Year/UpperSnake_GSI_SectionYear_output.RDS")

# spread data
gsi_bscorr <- gsi_dat$bootstrapped_proportions %>%  
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>% 
  mutate(bscorrnum = bs_corrected_repunit_ppn, section = str_sub(mixture_collection, end = -6), year = str_sub(mixture_collection, start = -4)) %>%
  select(section, year, repunit, bscorrnum) %>% spread(key = repunit, value = bscorrnum)

# separate grouping variables 
gsi_bscorr_grp <- gsi_bscorr[,c(1:2)]
gsi_bscorr_grp_fact <- gsi_bscorr_grp %>% mutate(section = recode(section, "snake_pacificdeadmans" = 1, "snake_deadmansmoose" = 2, "snake_moosewilson" = 3, "snake_wilsonsouthpark" = 4, "snake_southparkastoria" = 5, "snake_astoriawesttable" = 6),
                                                 year = as.numeric(year)-2019)

# separate bootstrapped proportions
gsi_bscorr_ppn <- gsi_bscorr[,-c(1:2)]
gsi_bscorr_ppn_fudge <- (gsi_bscorr_ppn + 0.0000001) / rowSums(gsi_bscorr_ppn + 0.0000001)
(gsi_bscorr_ppn)
```


### Bootstrapped mixing counts

Multinomial takes integer count data, so multiply bootstrapped proportions by collection sample size and round to the nearest integer to generate "bootstrapped counts".
```{r}
gsi_bscorr_abun <- gsi_dat$bootstrapped_proportions %>%
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>%
  mutate(bscorrnum = bs_corrected_repunit_ppn*sampsize, section = str_sub(mixture_collection, end = -6), year = str_sub(mixture_collection, start = -4)) %>%
  arrange(mixture_collection, repunit) %>% select(mixture_collection, section, year, repunit, bscorrnum) %>% spread(key = repunit, value = bscorrnum)
gsi_bscorr_abun <- gsi_bscorr_abun[,-c(1:3)]
gsi_bscorr_abun_int <- round(gsi_bscorr_abun, digits = 0)
(gsi_bscorr_abun_int)
```

### Covariate data

Groundwater, distance, and basin size
```{r}
# groundwater metrics
gwmet <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Groundwater/GroundwaterMetrics_raw_RepUnits.csv") %>% 
  rename(repunit = site) %>% 
  mutate(logarea = log(areasqkm), loggwi = log(gwi_iew05km)) %>% 
  mutate(z_logarea = as.numeric(scale(logarea)), z_loggwi = as.numeric(scale(loggwi))) %>%
  arrange(repunit)
# gwmet$repunit <- str_replace_all(gwmet$repunit, "deadman _greys", "deadman_greys")

# flowline distances
dist <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Flowline Distance/SnakeRiverSections_RepUnits_FlowlineDistance.csv") %>%
  mutate(z_dist = as.numeric(scale(distkm)))

# barriers
barr <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Barriers/RepUnits_BarrierSummary.csv") %>%
  mutate(log_barrier_dens = log(barrier_dens + 0.0001),
         z_barrier_dens = as.numeric(scale(barrier_dens)),
         z_barrier_num = as.numeric(scale(numbarr)),
         z_log_barrier_dens = as.numeric(scale(log_barrier_dens)),
         connect_type = as.factor(ifelse(is.na(connect_type), "connected", connect_type))) %>%
  mutate(connect_type = recode(connect_type, "culvert" = "culvert_diversion", "diversion dam" = "culvert_diversion"))

# Landcover
land <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Landcover/RepUnits_LandcoverSummary.csv") %>%
  mutate(log_prop_ag = log(prop_ag + 0.0001),
         log_prop_dev = log(prop_dev + 0.0001),
         log_prop_agdev = log(prop_agdev + 0.0001),
         z_prop_ag = as.numeric(scale(prop_ag)),
         z_prop_dev = as.numeric(scale(prop_dev)),
         z_prop_agdev = as.numeric(scale(prop_agdev)),
         z_log_prop_ag = as.numeric(scale(log_prop_ag)),
         z_log_prop_dev = as.numeric(scale(log_prop_dev)),
         z_log_prop_agdev = as.numeric(scale(log_prop_agdev)))

# scale tibble
scltib <- tibble(param = c("logarea", "loggwi", "dist", "barrier_dens", "prop_ag", "prop_dev", "prop_agdev"), 
                 mean = c(attributes(scale(gwmet$logarea))$`scaled:center`, attributes(scale(gwmet$loggwi))$`scaled:center`, attributes(scale(dist$distkm))$`scaled:center`, attributes(scale(barr$barrier_dens))$`scaled:center`, attributes(scale(land$prop_ag))$`scaled:center`, attributes(scale(land$prop_dev))$`scaled:center`, attributes(scale(land$prop_agdev))$`scaled:center`),  
                 sd = c(attributes(scale(gwmet$logarea))$`scaled:scale`, attributes(scale(gwmet$loggwi))$`scaled:scale`, attributes(scale(dist$distkm))$`scaled:scale`, attributes(scale(barr$barrier_dens))$`scaled:scale`, attributes(scale(land$prop_ag))$`scaled:scale`, attributes(scale(land$prop_dev))$`scaled:scale`, attributes(scale(land$prop_agdev))$`scaled:scale`))
write_csv(scltib, "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/GSI_Covariates_MeansStDevs.csv")

# covariates for JAGS
dat <- read_csv("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/GSI Analysis/By Section and Year/UpperSnake_GSI_SectionYear_BootstrappedProportions.csv") %>% 
  filter(mixture_collection != "snake_jldcattlemens") %>% 
  rename(mixture = mixture_collection, ppn_bscor = bs_corrected_repunit_ppn) %>%
  mutate(section = str_sub(mixture, end = -6), year = as.numeric(str_sub(mixture, start = -4))) %>%
  arrange(mixture, repunit)

dist_cov <- dat %>% left_join(dist) %>% arrange(mixture, repunit) %>% select(mixture, repunit, z_dist) %>% spread(key = repunit, value = z_dist) %>% select(-mixture)
area_cov <- c(gwmet$z_logarea)
gw_cov <- c(gwmet$z_loggwi)
```

### Create long data
```{r}
data_long <- gsi_dat$bootstrapped_proportions %>%  
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>% 
  mutate(bscorrnum = round(bs_corrected_repunit_ppn*sampsize, digits = 0),
         section = str_sub(mixture_collection, end = -6), 
         year = str_sub(mixture_collection, start = -4)) %>% 
  left_join(gwmet) %>% left_join(dist) %>% left_join(barr %>% select(repunit, barrier_dens, log_barrier_dens, connectivity, z_barrier_dens, z_log_barrier_dens, z_barrier_num, connect_type)) %>% left_join(land)
(data_long)
```


## Explore data

How correlated are the covariates?
```{r fig.width=8, fig.height=8}
ggpairs(data_long %>% select(z_dist, z_logarea, z_loggwi, connectivity, z_barrier_num, z_log_prop_agdev))
```

Plot proportional contribution against covariates
```{r fig.width=7, fig.height=9}
p1 <- data_long %>% ggplot(aes(x = z_dist, y = bs_corrected_repunit_ppn)) + geom_point() + geom_smooth()
p2 <- data_long %>% ggplot(aes(x = z_logarea, y = bs_corrected_repunit_ppn)) + geom_point() + geom_smooth()
p3 <- data_long %>% ggplot(aes(x = z_loggwi, y = bs_corrected_repunit_ppn)) + geom_point() + geom_smooth()
p4 <- data_long %>% ggplot(aes(x = as.factor(connectivity), y = bs_corrected_repunit_ppn)) + geom_boxplot() + geom_smooth()
p5 <- data_long %>% ggplot(aes(x = z_barrier_dens, y = bs_corrected_repunit_ppn)) + geom_point() + geom_smooth()
p6 <- data_long %>% ggplot(aes(x = z_prop_agdev, y = bs_corrected_repunit_ppn)) + geom_point() + geom_smooth(se = F)
ggarrange(p1, p2, p3, p4, p5, p6, nrow = 3, ncol = 2)
```

```{r fig.width=5, fig.height=4}
data_long %>% filter(z_loggwi >0) %>% ggplot(aes(x = z_dist, y = bs_corrected_repunit_ppn)) + geom_point() + ylim(0, 0.5)
data_long %>% filter(z_loggwi <= 0) %>% ggplot(aes(x = z_dist, y = bs_corrected_repunit_ppn)) + geom_point() + ylim(0, 0.5)

data_long %>% ggplot(aes(x = z_dist, y = bs_corrected_repunit_ppn, color = z_logarea)) + geom_point()
data_long %>% ggplot(aes(x = z_dist, y = bs_corrected_repunit_ppn, color = z_loggwi)) + geom_point()
data_long %>% ggplot(aes(x = connect_type, y = bs_corrected_repunit_ppn)) + geom_boxplot()
```

```{r}
barr %>% group_by(connect_type) %>% summarize(n.streams = n()) %>% left_join(data_long %>% group_by(connect_type) %>% summarize(n.obs = n()))
```



On the logit scale
```{r fig.width=7, fig.height=9}
logit <- function (x) {log(x/(1-x))}
data_long <- data_long %>% mutate(logitppn = logit(bs_corrected_repunit_ppn+0.001))

p1 <- data_long %>% ggplot(aes(x = z_dist, y = logitppn)) + geom_point()
p2 <- data_long %>% ggplot(aes(x = z_logarea, y = logitppn)) + geom_point()
p3 <- data_long %>% ggplot(aes(x = z_loggwi, y = logitppn)) + geom_point()
p4 <- data_long %>% ggplot(aes(x = as.factor(connectivity), y = logitppn)) + geom_boxplot()
p5 <- data_long %>% ggplot(aes(x = z_barrier_dens, y = logitppn)) + geom_point()
p6 <- data_long %>% ggplot(aes(x = z_prop_agdev, y = logitppn)) + geom_point()
ggarrange(p1, p2, p3, p4, p5, p6, nrow = 3, ncol = 2)
```









## Dirichlet-multinomial regression

This was the original model parameterization, as in the dissertation.

### Declare model
```{r}
cat("model {

##-----------LIKELIHOOD-------------------##
for (i in 1:Nmix) {
    Y[i,1:Nrgs] ~ dmulti(p[i,1:Nrgs], N[i])
    for (j in 1:Nrgs) {
        p[i,j] <- delta[i,j] / sum(delta[i,])
        delta[i,j] ~ dgamma(a[i,j], 1) T(0.0001,)
        log(a[i,j]) <- beta0 + beta1*dist[i,j] + beta2*area[j] + beta3*gw[j]
        loglik[i,j] <- logdensity.gamma(delta[i,j], a[i,j], 1) + logdensity.multi(Y[i,1:Nrgs], p[i,1:Nrgs], N[i])
    }
}

##-----------PRIORS-----------------------##

# INTERCEPT
beta0 ~ dnorm(0, pow(10, -2))

# for (i in 1:Nmix) { phi[i] ~ dgamma(phi_shape, phi_rate) }
# phi_shape ~ dgamma(1,1)
# phi_rate ~ dgamma(1,1)

# SLOPES
beta1 ~ dnorm(0, pow(10, -2))
beta2 ~ dnorm(0, pow(10, -2))
beta3 ~ dnorm(0, pow(10, -2))

# # INTENSITY
# for (i in 1:Nmix) {
#     theta[i] ~ dunif(0, 4000)
#     }

##-----------DERIVED VALUES-----------------------##
for (i in 1:Nmix) {
    y_pred2[i,1:Nrgs] ~ dmulti(p[i,1:Nrgs], N[i])
    y_pred[i,1:Nrgs] ~ dmulti(p_pred[i,1:Nrgs], N[i])
    for (j in 1:Nrgs) {
        p_pred[i,j] <- delta_pred[i,j] / sum(delta_pred[i,])
        delta_pred[i,j] ~ dgamma(a[i,j], 1) T(0.0001,)
    }
}

}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/MultiDirichMod_Covars.txt")
```

### Fit the model
```{r}
# specify data
jags.data <- list("Nmix" = dim(gsi_bscorr_ppn_fudge)[1], "Nrgs" = dim(gsi_bscorr_ppn_fudge)[2], 
                  "Y" = gsi_bscorr_abun_int, "N" = rowSums(gsi_bscorr_abun_int),
                  "dist" = dist_cov, "area" = area_cov, "gw" = gw_cov)

# parameters to monitor
params <- c("a", "beta0", "p_pred", "y_pred", "y_pred2", "loglik", "Y", "p", "beta1", "beta2", "beta3")   # parameters to monitor

# model file 
mymodel <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/MultiDirichMod_Covars.txt"

# run in JAGS
mod_md <- jags.parallel(jags.data, inits = NULL, parameters.to.save = params, model.file = mymodel, n.chains = 3, n.thin = 50, n.burnin = 2000, n.iter = 17000, DIC = TRUE)
```

Get MCMC samples and summary
```{r}
top_mod <- mod_md
# generate MCMC samples and store as an array
modelout <- top_mod$BUGSoutput
McmcList <- vector("list", length = dim(modelout$sims.array)[2])
for(i in 1:length(McmcList)) { McmcList[[i]] = as.mcmc(modelout$sims.array[,i,]) }
# rbind MCMC samples from 10 chains 
Mcmcdat <- rbind(McmcList[[1]], McmcList[[2]], McmcList[[3]])
param.summary <- modelout$summary
head(param.summary)
```


### Model diagnostics

#### R-hat

Any problematic R-hat values (>1.1)?
```{r}
mod_md$BUGSoutput$summary[,8][mod_md$BUGSoutput$summary[,8] > 1.1]
```

#### Traceplots

Ensure stationarity and sufficient mixing among chains
```{r}
MCMCtrace(mod_md, ind = TRUE, params = c("beta0", "beta1", "beta2", "beta3"), pdf = FALSE)
```


#### LOO-CV

Pareto k diagnostic values indicate the model is not well-specified:
```{r}
# get log-likelihoods
loglik_m <- t(apply(mod_md$BUGSoutput$sims.list$loglik, 1, c))

# get relative effective sample size
reff_m <- relative_eff(exp(loglik_m), chain_id = c(rep(1,300), rep(2,300), rep(3,300)))

# calculate loo
loo_m <- loo(loglik_m, r_eff = reff_m)

# compare
print(loo_m)
plot(loo_m)
```

#### Bayesian p-value

Get observed and expected values
```{r}
ppdat_obs <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y")])
ppdat_exp <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "y_pred[")])
```

Bayesian p-value << 0.50 indicates model predictions are biased
```{r}
sum(ppdat_exp > ppdat_obs) / (dim(ppdat_obs)[1]*dim(ppdat_obs)[2])
```


#### PP check

Graphical posterior predictive check indicates the model cannot make accurate predictions
```{r}
par(mar = c(4.5,4.5,1,1))
plot(jitter(apply(ppdat_exp, 2, median)) ~ jitter(apply(ppdat_obs, 2, median)), xlab = "Observed", ylab = "Expected")
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```

Previously, I failed to incorporate uncertainty from the Dirichlet process into the PP-check, which yielded the following graphical check, inflating our confidence in the model
```{r}
ppdat_obs <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y")])
ppdat_exp <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "y_pred2[")])
par(mar = c(4.5,4.5,1,1))
plot(jitter(apply(ppdat_exp, 2, median)) ~ jitter(apply(ppdat_obs, 2, median)), xlab = "Observed", ylab = "Expected")
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```


## Dirichlet regression

The parameterization forgoes the multinomial component, using transformed bootstrap-corrected mixing proportions. I.e., transformed 0s to very small numbers (b/c Dirichlet regression cannot handle 0s) and recalculated proportions to ensure sume to 1 constraint. Note that transformation can lead to bias model output (Douma and Weedon 2019). Also note that in forgoing the multinomial specificiation the model has no information on sample size/sampling effort...although this is probably not important as sample density was ~constant across mixtures.

### Declare model
```{r eval=FALSE, include=FALSE}
cat("model {

##-----------LIKELIHOOD-------------------##
for (i in 1:Nmix) {
    Y[i,1:Nrgs] ~ ddirch(a[i,1:Nrgs])
    loglik[i] <- logdensity.dirch(Y[i,1:Nrgs], a[i,1:Nrgs])
    for (j in 1:Nrgs) {
        log(a[i,j]) <- beta0[section[i], year[i],j]
    }
}

##-----------PRIORS-----------------------##

for (s in 1:Nsect) {
    for (y in 1:Nyear) {
        for (j in 1:Nrgs) {
            beta0[s,y,j] ~ dnorm(0,0.01)
        }
    }
}

##-----------DERIVED VALUES-----------------------##
for (i in 1:Nmix) {
    for (j in 1:Nrgs) {
        pred.ppn[i,j] <- a[i,j] / sum(a[i,])
        res[i,j] <- Y[i,j] - pred.ppn[i,j]
    }
}

}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj1 Spatiotemporal Var/DirichMod_Space.txt")
```
```{r}
cat("model {

##-----------LIKELIHOOD-------------------##
for (i in 1:Nmix) {
    Y[i,1:Nrgs] ~ ddirch(a[i,1:Nrgs])
    loglik[i] <- logdensity.dirch(Y[i,1:Nrgs], a[i,1:Nrgs])
    for (j in 1:Nrgs) {
        log(a[i,j]) <- beta0 + beta1*dist[i,j] + beta2*area[j] + beta3*gw[j]
    }
}

##-----------PRIORS-----------------------##

# INTERCEPT
beta0 ~ dnorm(0, pow(10, -2))

# SLOPES
beta1 ~ dnorm(0, pow(10, -2))
beta2 ~ dnorm(0, pow(10, -2))
beta3 ~ dnorm(0, pow(10, -2))

# # INTENSITY
# for (i in 1:Nmix) {
#     theta[i] ~ dunif(0, 4000)
#     }

##-----------DERIVED VALUES-----------------------##
for (i in 1:Nmix) {
    p_pred[i,1:Nrgs] ~ ddirch(a[i,1:Nrgs])
}

}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/DirichMod_Covars.txt")
```

### Fit the model
```{r}
# specify data
jags.data <- list("Nmix" = dim(gsi_bscorr_ppn_fudge)[1], "Nrgs" = dim(gsi_bscorr_ppn_fudge)[2], 
                  "Y" = gsi_bscorr_ppn_fudge,
                  "dist" = dist_cov, "area" = area_cov, "gw" = gw_cov)

# parameters to monitor
params <- c("a", "beta0", "p_pred", "loglik", "Y", "beta1", "beta2", "beta3")   # parameters to monitor

# model file 
mymodel <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/DirichMod_Covars.txt"

# run in JAGS
mod_dir <- jags.parallel(jags.data, inits = NULL, parameters.to.save = params, model.file = mymodel, n.chains = 3, n.thin = 50, n.burnin = 2000, n.iter = 17000, DIC = TRUE)
```

Get MCMC samples and summary
```{r}
top_mod <- mod_dir
# generate MCMC samples and store as an array
modelout <- top_mod$BUGSoutput
McmcList <- vector("list", length = dim(modelout$sims.array)[2])
for(i in 1:length(McmcList)) { McmcList[[i]] = as.mcmc(modelout$sims.array[,i,]) }
# rbind MCMC samples from 10 chains 
Mcmcdat <- rbind(McmcList[[1]], McmcList[[2]], McmcList[[3]])
param.summary <- modelout$summary
head(param.summary)
```


### Model diagnostics

#### R-hat

Any problematic R-hat values (>1.05)?
```{r}
mod_dir$BUGSoutput$summary[,8][mod_dir$BUGSoutput$summary[,8] > 1.05]
```

#### Traceplots

Ensure stationarity and sufficient mixing among chains
```{r}
MCMCtrace(mod_dir, ind = TRUE, params = c("beta0", "beta1", "beta2", "beta3"), pdf = FALSE)
```

#### LOO-CV
```{r}
# get log-likelihoods
loglik_m <- mod_dir$BUGSoutput$sims.list$loglik

# get relative effective sample size
reff_m <- relative_eff(exp(loglik_m), chain_id = c(rep(1,300), rep(2,300), rep(3,300)))

# calculate loo
loo_m <- loo(loglik_m, r_eff = reff_m)

# compare
print(loo_m)
plot(loo_m)
```


#### Bayesian p-value

Get observed and expected values
```{r}
ppdat_obs <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y")])
ppdat_exp <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "p_pred")])
```

Bayesian p-value suggests little bias in predictions, but this is misleading giving graphical pp-check
```{r}
sum(ppdat_exp > ppdat_obs) / (dim(ppdat_obs)[1]*dim(ppdat_obs)[2])
```


#### PP check

Graphical posterior predictive check indicates considerable issues with accuracy and precision of model predictions
```{r}
par(mar = c(4.5,4.5,1,1))
plot(apply(ppdat_exp, 2, median) ~ apply(ppdat_obs, 2, median), xlab = "Observed", ylab = "Expected")
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```

Same as above but with axis limits set to be equal
```{r}
par(mar = c(4.5,4.5,1,1))
plot(apply(ppdat_exp, 2, median) ~ apply(ppdat_obs, 2, median), xlab = "Observed", ylab = "Expected", xlim = c(0, 0.5), ylim = c(0, 0.5))
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```


## Multinomial regression

What if we just try a multinomial regression? The prediction was bad for Dirichlet-multinomial model and worse for the Dirichlet only model...could the problem be in the Dirichlet process?

### Declare model
```{r}
cat("model {

##-----------LIKELIHOOD-------------------##
for (i in 1:Nmix) {
    Y[i,1:Nrgs] ~ dmulti(p[i,1:Nrgs], N[i])
    for (j in 1:Nrgs) {
        logit(p[i,j]) <- beta0 + beta1*dist[i,j] + beta2*area[j] + beta3*gw[j]
    }
    loglik[i] <- logdensity.multi(Y[i,1:Nrgs], p[i,1:Nrgs], N[i])
}

##-----------PRIORS-----------------------##

# INTERCEPT
beta0 ~ dnorm(0, pow(10, -2))

# SLOPES
beta1 ~ dnorm(0, pow(10, -2))
beta2 ~ dnorm(0, pow(10, -2))
beta3 ~ dnorm(0, pow(10, -2))

##-----------DERIVED VALUES-----------------------##
for (i in 1:Nmix) {
    y_pred[i,1:Nrgs] ~ dmulti(p[i,1:Nrgs], N[i])
}

}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/MultiMod_Covars.txt")
```

### Fit the model
```{r}
# specify data
jags.data <- list("Nmix" = dim(gsi_bscorr_ppn_fudge)[1], "Nrgs" = dim(gsi_bscorr_ppn_fudge)[2], 
                  "Y" = gsi_bscorr_abun_int, "N" = rowSums(gsi_bscorr_abun_int),
                  "dist" = dist_cov, "area" = area_cov, "gw" = gw_cov)

# parameters to monitor
params <- c("beta0", "y_pred", "loglik", "Y", "p", "beta1", "beta2", "beta3")   # parameters to monitor

# model file 
mymodel <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/MultiMod_Covars.txt"

# run in JAGS
mod_mult <- jags.parallel(jags.data, inits = NULL, parameters.to.save = params, model.file = mymodel, n.chains = 3, n.thin = 50, n.burnin = 2000, n.iter = 17000, DIC = TRUE)
```

Get MCMC samples and summary
```{r}
top_mod <- mod_mult
# generate MCMC samples and store as an array
modelout <- top_mod$BUGSoutput
McmcList <- vector("list", length = dim(modelout$sims.array)[2])
for(i in 1:length(McmcList)) { McmcList[[i]] = as.mcmc(modelout$sims.array[,i,]) }
# rbind MCMC samples from 10 chains 
Mcmcdat <- rbind(McmcList[[1]], McmcList[[2]], McmcList[[3]])
param.summary <- modelout$summary
head(param.summary)
```


### Model diagnostics

#### R-hat

Any problematic R-hat values (>1.1)?
```{r}
mod_mult$BUGSoutput$summary[,8][mod_mult$BUGSoutput$summary[,8] > 1.1]
```

#### Traceplots

Ensure stationarity and sufficient mixing among chains
```{r}
MCMCtrace(mod_mult, ind = TRUE, params = c("beta0", "beta1", "beta2", "beta3"), pdf = FALSE)
```


#### LOO-CV

Pareto k diagnostic values indicate the model is not well-specified:
```{r}
# get log-likelihoods
loglik_m <- mod_mult$BUGSoutput$sims.list$loglik

# get relative effective sample size
reff_m <- relative_eff(exp(loglik_m), chain_id = c(rep(1,300), rep(2,300), rep(3,300)))

# calculate loo
loo_m <- loo(loglik_m, r_eff = reff_m)

# compare
print(loo_m)
plot(loo_m)
```

#### Bayesian p-value

Get observed and expected values
```{r}
ppdat_obs <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y")])
ppdat_exp <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "y_pred[")])
```

Bayesian p-value << 0.50 indicates model predictions are biased
```{r}
sum(ppdat_exp > ppdat_obs) / (dim(ppdat_obs)[1]*dim(ppdat_obs)[2])
```


#### PP check

Graphical posterior predictive check indicates the model cannot make accurate predictions
```{r}
par(mar = c(4.5,4.5,1,1))
plot(jitter(apply(ppdat_exp, 2, median)) ~ jitter(apply(ppdat_obs, 2, median)), xlab = "Observed", ylab = "Expected")
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```




## Zero-inflated beta regr.

Seems like the multivariate options may not be the most appropriate approach given that we are interested in the effects of covariates at the scale of 

### Fit model
```{r}
# data_long <- data_long %>% mutate(connectivity_fac = as.factor())

mod_zoib1 <- brm(brmsformula(
  bs_corrected_repunit_ppn ~ z_dist + z_logarea + z_loggwi + z_log_prop_agdev + connect_type, 
  phi ~ z_dist + z_logarea + z_loggwi + z_log_prop_agdev + connect_type, 
  zi ~ z_dist + z_logarea + z_loggwi + z_log_prop_agdev + connect_type, 
  family = zero_inflated_beta()), data = data_long, iter = 2000)

#connectivity + z_dist*connectivity + z_logarea*connectivity + z_loggwi*connectivity.  z_dist*z_logarea + z_dist*z_loggwi 
```

```{r}
pairs(mod_zoib1)
```


```{r}
summary(mod_zoib1)
```


```{r fig.width=7, fig.height=7}
plot(mod_zoib1, ask = FALSE)
```



```{r}
bayes_R2(mod_zoib1)
```


```{r}
yrep <- brms::posterior_predict(mod_zoib1, ndraws = 500)
dim(yrep)
```

```{r}
yrep_sum <- apply(yrep, 2, median)
length(which(yrep_sum > data_long$bs_corrected_repunit_ppn)) / length(data_long$bs_corrected_repunit_ppn)
```


```{r}
myloo <- loo(mod_zoib1)
myloo
plot(myloo, label_points = TRUE)
```

```{r}
pp_check(mod_zoib1, type = "hist", ndraws = 5)
pp_check(mod_zoib1, type = "error_hist", ndraws = 5) + geom_vline(xintercept = 0, linetype = "dashed") + xlim(-0.2, 0.2)
ppc_stat_2d(y = data_long$bs_corrected_repunit_ppn, yrep = yrep, stat = c("mean", "var"))
# ppc_dens_overlay(y = data_long$bs_corrected_repunit_ppn, yrep = yrep[1:100,])
# ppc_dens_overlay(y = data_long$bs_corrected_repunit_ppn, yrep = yrep[1:100,]) + coord_cartesian(xlim = c(0, .1))
# ppc_dens_overlay(y = data_long$bs_corrected_repunit_ppn, yrep = yrep[1:100,]) + xlim(.1, .50)
```

```{r}
ppc_scatter_avg(y = data_long$bs_corrected_repunit_ppn, yrep = yrep, alpha = 0.5) + coord_flip()

# yrep_sum <- apply(yrep, 2, median)
# plot(jitter(yrep_sum) ~ jitter(data_long$bscorrnum))
# abline(a = 0, b = 1, col = "red")
```

```{r, fig.height=12, fig.width=12}
ppc_scatter_avg_grouped(y = data_long$bs_corrected_repunit_ppn, yrep = yrep[1:100,], group = data_long$repunit) + coord_flip()

tibble(repunit = data_long$repunit, obs = data_long$bs_corrected_repunit_ppn, pred = yrep_sum) %>%
  ggplot(aes(x = obs, y = pred)) + geom_point() + geom_abline(intercept = 0, slope = 1, linetype = "dashed") + facet_wrap(~repunit)
```



```{r}
prop_zero <- function(x) mean(x == 0)

ppc_stat(y = data_long$bs_corrected_repunit_ppn, yrep = yrep, stat = "prop_zero")
ppc_stat(y = data_long$bs_corrected_repunit_ppn, yrep = yrep, stat = "mean")
# 
# ppc_stat(y = data_long$bscorrnum, yrep = yrep, stat = "prop_zero")
# ppc_stat(y = data_long$bscorrnum, yrep = yrep, stat = "max")
```


```{r}
plot(conditional_effects(mod_zoib1), ask = FALSE, rug = TRUE)
```

```{r}
plot(conditional_effects(mod_zoib1, dpar = "zi"), ask = FALSE)
```

```{r}
plot(conditional_effects(mod_zoib1, dpar = "phi"), ask = FALSE)
```


```{r}
pred_beta_zi <- mod_zoib1 %>% epred_draws(newdata = expand_grid(z_dist = 0, z_logarea = 0, z_loggwi = 0, connect_type = c("connected", "low flow")), dpar = "zi")

pred_beta_zi %>% group_by(connect_type) %>% median_hdci(zi)
pred_beta_zi %>% compare_levels(variable = zi, by = connect_type) %>% median_hdci()
```



## Binomial
```{r}
mod_zoib <- brm(bscorrnum | trials(sampsize) ~ (1|repunit + section + year) + z_dist + z_logarea + z_loggwi, family = binomial(link = "logit"), data = data_long, iter = 3000)
```

```{r}
binmod <- glm(bs_corrected_repunit_ppn + 0.0001 ~ z_dist + z_logarea + z_loggwi + connect_type, data = data_long, family = quasibinomial())

# binmod <- glm(cbind(bscorrnum, sampsize-bscorrnum) ~ z_dist + z_logarea + z_loggwi + connect_type, data = data_long, family = quasibinomial())


# Calculate R-squared
deviance <- summary(binmod)$deviance
null_deviance <- summary(binmod)$null.deviance
rsquared <- 1 - (deviance / null_deviance)

# Print the R-squared value
print(rsquared)
summary(binmod)$r.squared

binpred <- predict(binmod, type = "response")
plot(binpred ~ data_long$bs_corrected_repunit_ppn)
```

```{r}
ddd <- data_long %>% mutate(bbb = bs_corrected_repunit_ppn+0.0001)

br <- betareg::betareg(bbb ~ z_dist + z_logarea + z_loggwi + connect_type, data = ddd, link = "logit")
summary(br)
br$pseudo.r.squared
brpred <- betareg::predict(br, type = "response")
plot(brpred ~ ddd$bbb)
```



```{r}
mod_beta <- brm(bf(bs_corrected_repunit_ppn+0.0001 ~ z_dist + z_logarea + z_loggwi + connect_type, family = Beta(link = "logit")), data = data_long, iter = 1000)
bayes_R2(mod_beta)
pp_check(mod_beta, "scatter_avg") + coord_flip()
```


```{r}
mod_binom <- brm(bf(bscorrnum | trials(sampsize) ~ z_dist + z_logarea + z_loggwi + connect_type, family = binomial()), data = data_long, iter = 1000)
bayes_R2(mod_binom)
pp_check(mod_binom, "scatter_avg") + coord_flip()
```
```{r}
loo_compare(loo(mod_beta), loo(mod_binom))
```




## Zero-inflated dirichlet regression

### Declare model
```{r}
cat("model {

##-----------LIKELIHOOD-------------------##
for (i in 1:Nmix) {
    for (j in 1:Nrgs) {
        # Zero-inflation component
        zero[i,j] ~ dbern(pi[j])
        y_nonzero[i,j] ~ ddirch(a[i, 1:Nrgs])

        # Observation model: combines zeros and Dirichlet
        Y[i,j] <- zero[i,j] * 0 + (1 - zero[i,j]) * y_nonzero[i,j]
        }

    # Linear model for alpha
    for (j in 1:Nrgs) {
        log(a[i,j]) <- beta0 + beta1*dist[i,j] + beta2*area[j] + beta3*gw[j]
        }
    }

##-----------PRIORS-----------------------##

# Priors for zero-inflation probabilities
for (j in 1:Nrgs) { pi[j] ~ dbeta(1, 1) }  # Uniform prior on [0, 1]

# INTERCEPT
beta0 ~ dnorm(0, pow(10, -2))

# SLOPES
beta1 ~ dnorm(0, pow(10, -2))
beta2 ~ dnorm(0, pow(10, -2))
beta3 ~ dnorm(0, pow(10, -2))

##-----------DERIVED VALUES-----------------------##
for (i in 1:Nmix) {
    for (j in 1:Nrgs) {
        y_pred[i,j] <- ddirch(a[i,j])
    }
}

}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/ZeroInfDirichMod_Covars.txt")
```

### Fit the model
```{r}
# specify data
jags.data <- list("Nmix" = dim(gsi_bscorr_ppn_fudge)[1], "Nrgs" = dim(gsi_bscorr_ppn_fudge)[2], 
                  # "Nsect" = length(unique(gsi_bscorr_grp_fact$section)), "section" = as.factor(gsi_bscorr_grp_fact$section), 
                  # "Nyear" = length(unique(gsi_bscorr_grp_fact$year)), "year" = as.factor(gsi_bscorr_grp_fact$year), 
                  "Y" = gsi_bscorr_ppn,
                  "dist" = dist_cov, "area" = area_cov, "gw" = gw_cov)

# parameters to monitor
params <- c("a", "beta0", "y_pred", "Y", "pi", "zero", "y_nonzero", "beta1", "beta2", "beta3")   # parameters to monitor

# model file 
mymodel <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/ZeroInfDirichMod_Covars.txt"

# run in JAGS
mod_zid <- jags.parallel(jags.data, inits = NULL, parameters.to.save = params, model.file = mymodel, n.chains = 3, n.thin = 20, n.burnin = 2000, n.iter = 12000, DIC = TRUE)
```

Get MCMC samples and summary
```{r}
top_mod <- mod_md
# generate MCMC samples and store as an array
modelout <- top_mod$BUGSoutput
McmcList <- vector("list", length = dim(modelout$sims.array)[2])
for(i in 1:length(McmcList)) { McmcList[[i]] = as.mcmc(modelout$sims.array[,i,]) }
# rbind MCMC samples from 10 chains 
Mcmcdat <- rbind(McmcList[[1]], McmcList[[2]], McmcList[[3]])
param.summary <- modelout$summary
head(param.summary)
```



## Beta regression

### Data
```{r}
dat <- gsi_dat$bootstrapped_proportions %>%  
  left_join(gsi_dat$indiv_posteriors %>% group_by(indiv, mixture_collection) %>% summarize(n = n()) %>% group_by(mixture_collection) %>% summarize(sampsize = n())) %>% 
  mutate(section = str_sub(mixture_collection, end = -6), year = str_sub(mixture_collection, start = -4)) %>%
  left_join(dist) %>% left_join(gwmet) 

# add small values to deal with 0s
dat <- dat %>% mutate(bs_corrected_repunit_ppn_dum = bs_corrected_repunit_ppn + min(dat$bs_corrected_repunit_ppn[dat$bs_corrected_repunit_ppn > 0]))

head(dat)
```

### Declare model
```{r}
cat("model {

##-----------LIKELIHOOD-------------------##

for (i in 1:Nobs) {
    Y[i] ~ dbeta(p[i], q[i])
    p[i] <- mu[i] * phi
    q[i] <- (1 - mu[i]) * phi
    logit(mu[i]) <- beta0 + beta1 * dist[i] + beta2 * area[i] + beta3 * gw[i]
}

##-----------PRIORS-----------------------##

# INTERCEPT
beta0 ~ dnorm(0, pow(10, -2))

# SLOPES
beta1 ~ dnorm(0, pow(10, -2))
beta2 ~ dnorm(0, pow(10, -2))
beta3 ~ dnorm(0, pow(10, -2))

phi ~ dgamma(0.1, 0.1)

##-----------DERIVED VALUES-----------------------##

for (i in 1:Nobs) { Y_pred[i] ~ dbeta(p[i], q[i]) }


}", file = "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/BetaRegMod_Covars.txt")
```


### Fit the model
```{r}
# specify data
jags.data <- list("Nobs" = dim(dat)[1], 
                  "Y" = dat$bs_corrected_repunit_ppn_dum, 
                  "dist" = dat$z_dist, "area" = dat$z_logarea, "gw" = dat$z_loggwi)

# parameters to monitor
params <- c("beta0", "beta1", "beta2", "beta3", "phi", "Y", "Y_pred", "mu")   # parameters to monitor

# model file 
mymodel <- "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Obj2 Model Contribution/BetaRegMod_Covars.txt"

# run in JAGS
mod_bet <- jags.parallel(jags.data, inits = NULL, parameters.to.save = params, model.file = mymodel, n.chains = 3, n.thin = 10, n.burnin = 1000, n.iter = 6000, DIC = TRUE)
```

Get MCMC samples and summary
```{r}
top_mod <- mod_bet
# generate MCMC samples and store as an array
modelout <- top_mod$BUGSoutput
McmcList <- vector("list", length = dim(modelout$sims.array)[2])
for(i in 1:length(McmcList)) { McmcList[[i]] = as.mcmc(modelout$sims.array[,i,]) }
# rbind MCMC samples from 10 chains 
Mcmcdat <- rbind(McmcList[[1]], McmcList[[2]], McmcList[[3]])
param.summary <- modelout$summary
head(param.summary)
```

### Model diagnostics

#### View R-hat

Any problematic R-hat values (>1.01)?
```{r}
mod_bet$BUGSoutput$summary[,8][mod_bet$BUGSoutput$summary[,8] > 1.01]
```

#### View traceplots

Ensure stationarity and sufficient mixing among chains
```{r}
MCMCtrace(mod_bet, ind = TRUE, params = c("beta0", "beta1", "beta2", "beta3", "phi"), pdf = FALSE)
```


#### PP check

Get observed and expected values
```{r}
ppdat_obs <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y[")])
ppdat_exp <- as.matrix(Mcmcdat[,startsWith(colnames(Mcmcdat), "Y_pred[")])
```

Bayesian p-value:
```{r}
sum(ppdat_exp > ppdat_obs) / (dim(ppdat_obs)[1]*dim(ppdat_obs)[2])
```


Graphical posterior predictive check:
```{r}
par(mar = c(4.5,4.5,1,1))
plot(apply(ppdat_exp, 2, median) ~ apply(ppdat_obs, 2, median), xlab = "Observed", ylab = "Expected")
abline(a = 0, b = 1, col = "red", lwd = 2)
legend("topleft", legend = "1:1", lwd = 2, col = "red", bty = "n")
```

