# Landcover

Purpose: derive zonal landscover statisitics for reporting unit watersheds

```{r include = FALSE}
library(tidyverse)
library(mapview)
library(terra)
library(knitr)
```


## Load spatial data

View 2020 National Landcover Dataset, access from [https://www.mrlc.gov/viewer/] on 20 January 2025
```{r fig.height=8, fig.width=5}
# basin
basin <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Data/Spatial/Basin Delineation/BasinDelineation/MajorBasins_Watersheds.shp") 
basin <- subset(basin, basin$site %in% c("SnakeHeadwaters"))

# 2020 landcover, access from https://www.mrlc.gov/viewer/ on 20 January 2025
nlcd <- rast("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Landcover/Annual_NLCD_LndCov_2020.tiff")
plot(nlcd)
nlcd <- project(nlcd, crs(basin))

# reporting unit watersheds
watersheds <- vect("/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Watershed Delineation/RepUnits_Watersheds.shp")
```


## Reclassify NLCD

Create reclassification functions:
```{r}
# Agricultural
agfun <- function(x) { ifelse(x %in% c(81:82), 1, 0) }

# Developed
devfun <- function(x) { ifelse(x %in% c(21:24),1, 0) }

# Agriculture and developed
ADfun <- function(x) { ifelse(x %in% c(21:24,81:82),1, 0) }
```

Reclassify NLCD
```{r fig.width = 8, fig.height = 3}
nlcd_ag <- app(nlcd, agfun)
nlcd_dev <- app(nlcd, devfun)
nlcd_AD <- app(nlcd, ADfun)

par(mfrow = c(1,3))
plot(nlcd_ag, main = "Agricultural")
plot(nlcd_dev, main = "Developed")
plot(nlcd_AD, main = "Ag + Dev")
```

### Zonal statistics

Compute zonal statistics (proportional cover) for each reporting unit watershed and combine into tibble
```{r}
nlcd_tib <- tibble(repunit = watersheds$site,
                   prop_ag = unlist(zonal(nlcd_ag, watersheds, fun = "mean")),
                   prop_dev = unlist(zonal(nlcd_dev, watersheds, fun = "mean")),
                   prop_agdev = unlist(zonal(nlcd_AD, watersheds, fun = "mean")))
nlcd_tib %>% kable()
```

Write to file
```{r}
write_csv(nlcd_tib, "/Users/jeffbaldock/Library/CloudStorage/GoogleDrive-jbaldock@uwyo.edu/Shared drives/wyo-coop-baldock/UWyoming/Snake River Cutthroat/Analyses/Snake River GSI Quarto/Landscape Covariates/Landcover/RepUnits_LandcoverSummary.csv")
```



